---
title: "A stroke prediction model"
output: html_document
author: "Karolina Dyrla-Mularczyk"
source: "Data from Coursera - Build and deploy a stroke prediction model using R"
---
```{r setup, include=FALSE}
library("knitr")
library("tidyverse")
library("plotly")
library("htmlwidgets")
source("Stroke_code.R")
```

## This is a report about prediction of stroke. Orginal data cames from Coursera Platform - "Build and deploy a stroke prediction model using R". The dataset includes information on patients such as:

- __age__,
- __BMI__,
- __average glucose level__,
- __heart disease__,
- __hypertension__
- __and more demographic details__

##The dataset consists of 5110 observations, with 249 patients experiencing a stroke. This imbalance indicates that only 4.87% of the total observations involve stroke occurrences.

***




```{r}
summary(stroke)
skim(stroke)
```

The average age in the study group is  
43.3. The average glucose level in this study group is 
106.



This chart shows .

```{r EU_FR=plot, dev='png', out.width="150%", echo=FALSE}
plot()
```

# Task Two: Build prediction models

```{r}

```




# Task Three: Evaluate and select prediction models

```{r}

```



# Task Four: Deploy the prediction model

```{r}

```




# Task Five: Findings and Conclusions
































