---
title: "A stroke prediction model"
output: html_document
author: "Karolina Dyrla-Mularczyk"
source: "Data from Coursera - Build and deploy a stroke prediction model using R"
---
```{r setup, include=FALSE}
library("knitr")
library("tidyverse")
library("plotly")
library("htmlwidgets")
source("Stroke_code.R")
```

## This is a report about prediction of stroke. Orginal data cames from Coursera Platform - "Build and deploy a stroke prediction model using R". The dataset includes information on patients such as:

- __age__,
- __BMI__,
- __average glucose level__,
- __heart disease__,
- __hypertension__
- __and more demographic details__

##The dataset consists of 5110 observations, with 249 patients experiencing a stroke. This imbalance indicates that only 4.87% of the total observations involve stroke occurrences.

***


```{r}
summary(stroke)
skim(stroke)
```

The average age in the study group is  
43.3. 
The average glucose level in this study group is 
106.15.
In the study group, one person marked the gender "other". The record was removed to make the data visualization more readable. Ultimately, 5,109 observations were obtained.

The charts show the distribution of demographic variables: gender, place of residence and place of work..

```{r gender=plot, dev='png', out.width="150%", echo=FALSE}
plot(gender)
```

```{r residence=plot, dev='png', out.width="150%", echo=FALSE}
plot(residence)
```

```{r work=plot, dev='png', out.width="150%", echo=FALSE}
plot(work)
```

# Task Two: Build prediction models

```{r}

```




# Task Three: Evaluate and select prediction models

```{r}

```



# Task Four: Deploy the prediction model

```{r}

```




# Task Five: Findings and Conclusions
































