---
title: "The Stroke Prediction Model"
author: "Karolina Dyrla-Mularczyk"
output: html_document

---
```{r setup, include=FALSE}
library("knitr")
library(dplyr)
library(skimr)
library(DataExplorer)
library(corrplot)
library(caret)
library(tidyverse)
library(tidyquant)
library(tidymodels)
library(ggplot2)
library(gganimate)
library(ggrepel)
library(ggdist)
library(ggthemes)
library(RColorBrewer)
library(knitr)
library(plotly)
library(htmlwidgets)
library(GGally)
library(stats)
library(coefplot)
library(rpart)
library(rpart.plot)
library(useful)
library(xgboost)
library(gmodels)
library(rstudioapi)
library(datasets) 
library(caTools)
library(party)
library(magrittr)
library(Boruta)
library(gridExtra)
library(randomForest)
library(DiagrammeR)
library(glmnet)
library(C50)
source("Stroke_code.R")
```

#This is a report about prediction of stroke. Original data came from Coursera's Course - "Build and deploy a stroke prediction model using R". 

Stroke is a leading cause of death and disability worldwide, with significant public health implications. Some key facts about stroke:

- __Stroke is the second leading cause of death globally, accounting for approximately 11% of all deaths (WHO).__
- __Around 80% of strokes are preventable through lifestyle modifications such as healthy diet, regular exercise, and avoiding tobacco use.__
- __High blood pressure is one of the most important modifiable risk factor for stroke.__
- __Early recognition of stroke symptoms and prompt medical intervention can significantly improve outcomes and reduce disability.__
- __Hence the need to identify risk factors and assess their importance in the stroke prediction model.This is important both from the perspective of the healthcare sector and from the patient's perspective.__

The dataset came from Kaggle via Coursera and includes information on patients such as:

- age,
- BMI,
- average glucose level,
- heart disease,
- hypertension,
- and more demographic details

The dataset consists of 5110 observations, with 249 patients experiencing a stroke. This imbalance indicates that only 4.87% of the total observations involve stroke occurrences.

***

A few general informations about dataset:

```{r}

dim(stroke) 
head(stroke) 
str(stroke) 
names(stroke)
summary(stroke)
skim(stroke)
```

The average age in the study group is  
__43.3__. 
The average glucose level in this study group is 
__106.15__.
In the study group, one person marked the gender "other". The record was removed to make the data visualization more readable. Ultimately, 5,109 observations were obtained.

Missing values were found in the "bmi" column. It was decided to enter the average of all BMI measurements in the missing fields and round it to two decimal places.

The charts below show the distribution of demographic variables: gender, place of residence and place of work.

```{r gender=plot, dev='png', out.width="150%", echo=FALSE}
plot(gender)
```

```{r residence=plot, dev='png', out.width="150%", echo=FALSE}
plot(residence)
```

```{r work=plot, dev='png', out.width="150%", echo=FALSE}
plot(work)
```

The charts below show the distribution of health related variables: hypertension, heart disease, smoking status.

```{r smoking=plot, dev='png', out.width="150%", echo=FALSE}
plot(smoking)
```

```{r hypertension=plot, dev='png', out.width="150%", echo=FALSE}
plot(hypertension)
```

```{r heart_disease=plot, dev='png', out.width="150%", echo=FALSE}
plot(heart_disease)
```

BMI data was organized according to the "BMI Classification Percentile And Cut Off Points" classification. The BMI criteria distributed in the study group are presented in the chart below.

```{r echo=FALSE, warning=FALSE, bmi_plot=plot, dev='png', out.width="150%"}
plot(bmi_plot)
```

As we could see, the majorty patients have overweight. This corresponds to disturbing reports from other studies regarding overweight. As we could read on the WHO page (https://www.who.int/news-room/fact-sheets/detail/obesity-and-overweight): 

>"In 2022, 2.5 billion adults (18 years and older) were overweight. Of these, 890 million were living with obesity. In 2022, 43% of adults aged 18 years and over were overweight and 16% were living with obesity.

In the studied group there is 1610 patients with overweight. Which constitutes __'percentage'__ of the entire group.

The chart below shows the distribution of patients with and without stroke depending on their smoking status.

```{r stroke_smok_stat, echo=FALSE}
kable(stroke_smok_stat, col.names = c("Patients without stroke", "Patients with stroke"))
```


#Conclusions
#Future studies
Future research should take into account more factors related to health behaviors, such as eating habits, physical activity, and alcohol consumption. In the future, clinical indicators as: waist-height ratio, visceral adipose issue, triglyceride-glucose index should also be considered. The use of blood thinning medications should also be considered. 