---
title: "A stroke prediction model"
author: "Karolina Dyrla-Mularczyk"
output:
  html_document: default
  pdf_document: default
  word_document: default
source: "Data from Coursera - Build and deploy a stroke prediction model using R"
---
```{r setup, include=FALSE}
library("knitr")
library(dplyr)
library(skimr)
library(DataExplorer)
library(corrplot)
library(caret)
library(tidyverse)
library(tidyquant)
library(tidymodels)
library(ggplot2)
library(gganimate)
library(ggrepel)
library(ggdist)
library(ggthemes)
library(RColorBrewer)
library(knitr)
library(plotly)
library(htmlwidgets)
library(GGally)
library(stats)
library(coefplot)
library(rpart)
library(rpart.plot)
library(useful)
library(xgboost)
library(gmodels)
library(rstudioapi)
library(datasets) 
library(caTools)
library(party)
library(magrittr)
library(Boruta)
library(gridExtra)
library(randomForest)
library(DiagrammeR)
library(glmnet)
library(C50)
source("Stroke_code.R")
```

This is a report about prediction of stroke. Original data came from Coursera Platform - "Build and deploy a stroke prediction model using R". The dataset includes information on patients such as:

- __age__,
- __BMI__,
- __average glucose level__,
- __heart disease__,
- __hypertension__,
- __and more demographic details__

The dataset consists of 5110 observations, with 249 patients experiencing a stroke. This imbalance indicates that only 4.87% of the total observations involve stroke occurrences.

***


```{r}
summary(stroke)
skim(stroke)
```

The average age in the study group is  
43.3. 
The average glucose level in this study group is 
106.15.
In the study group, one person marked the gender "other". The record was removed to make the data visualization more readable. Ultimately, 5,109 observations were obtained.

The charts show the distribution of demographic variables: gender, place of residence and place of work.

```{r gender=plot, dev='png', out.width="150%", echo=FALSE}
plot(gender)
```

```{r residence=plot, dev='png', out.width="150%", echo=FALSE}
plot(residence)
```

```{r work=plot, dev='png', out.width="150%", echo=FALSE}
plot(work)
```

##The charts show the distribution of health related variables: hypertension, heart disease, smoking status.







